---
title-meta: FINA5250 Assignment 4
author-meta: LIU, Weiyang
geometry: "left=1.25in,right=1.25in,top=1in,bottom=1in"
output: pdf_document
---

# Assignment 4
FINA 5250 Empirical Methods in Finance

LIU, Weiyang

[wliuax@connect.ust.hk](mailto:wliuax@connect.ust.hk)

```{r setup, include = TRUE, echo = FALSE}
# set working directory
setwd("d:/workspace/FINA5250/Assignment4")
# remove all objects from the current workspace
rm(list = ls())
```

## Question 1
```{r, echo = FALSE}
cov_matrix <- function(sd_m, betas, taus = numeric()) {
  n <- length(betas)
  if (length(taus) == 0) {
    taus <- numeric(length = n)
  } else if (length(taus) != n) {
    cat("Error: Number of betas and taus do not match.\n")
    return(NA)
  }

  cm <- matrix(nrow = n, ncol = n)
  for (i in seq_len(n)) {
    for (j in seq_len(n)) {
      cm[i, j] <- betas[i] * betas[j] * sd_m**2
      if (i == j) {
        cm[i, j] <- cm[i, j] + taus[i]**2
      }
    }
  }

  return(cm)
}
```

### 1.1
The covariance matrix $\Sigma_{t+1}$ is as follows:
```{r, echo = FALSE}
sd_m <- 0.02
betas <- c(0.5, 1.5, 1.2)

cm <- cov_matrix(sd_m, betas)
print(cm)
```

### 1.2
Assume $R_{P,t+1}$ is standard normal, then the 1\% VaR for the one-day ahead return of B is:
```{r, echo = FALSE}
- qnorm(0.01) * sqrt(cm[2, 2])
```

### 1.3
Assume $R_{P,t+1}$ is standard normal, then the 1\% VaR for the one-day ahead return of the portfolio is:
```{r, echo = FALSE}
w <- c(0.5, 0.25, 0.25)
var_p <- t(w) %*% cm %*% w
- qnorm(0.01) * sqrt(var_p)[, ]
```
The beta of the portfolio is:
```{r, echo = FALSE}
(t(betas) %*% w)[, ]
```

## Question 2
```{r, echo = FALSE}
data <- read.table("Assignment4_Q2.csv", sep = ",", header = TRUE)
ret <- diff(log(as.matrix(data[c("DJI", "Gold", "USD.JPY")])))
ret_means <- colMeans(ret)
ret_cov <- cov(ret)

w <- rep(1 / 3, times = 3)
p_mean <- (ret_means %*% w)[, ]
p_var <- t(w) %*% ret_cov %*% w
p_sd <- sqrt(p_var)[, ]
```
The 1-day, 1\% VaR of of the portfolio is:
```{r, echo = FALSE}
- qnorm(0.01, p_mean, p_sd)
```

The 1-day, 1\% VaR of of DJIA is:
```{r, echo = FALSE}
- qnorm(0.01, ret_means[1], sqrt(ret_cov[1, 1]))
```

The VaR of the portfolio is lower than that of DJIA, because part of the unsystematic risk is diversified.
